---
title: "Fragile Families - Data Cleaning and Subsetting"
author: "Havi Khurana"
date: "5/29/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      error = FALSE,
                      warning = FALSE)
```


```{r}
library(tidyverse)
library(here)
library(janitor)
library(stringr)
```


```{r}
#importing meta-data
ff_meta <- read.csv(here("data","FFMetadata_v09 (1).csv"))%>%
   clean_names() %>%
  as_tibble()
```

Scales to clean:
- Child Behavior Problems (CBCL)
- Self-Description Questionnaire (SDQ)
- Teen Tobacco Use
- Peer Bullying 
- Trouble at School


```{r}
#select scales
scales_multiple <- c("Child Behavior Problems (CBCL)",
                     "Peer Bullying")
scales_once <- c("Self-Description Questionnaire (SDQ)",
            "Teen Tobacco Use",
            "Trouble at School")
```


```{r}
scale_vars <- ff_meta %>% 
                filter(scale %in% scales_once |
                    (scale %in% scales_multiple &
                     wave == "Year 15")) %>%
                select(new_name,
                       varlab,
                       scale,
                       wave,
                       respondent,
                       survey)

```

```{r, eval = FASE}
#meta-data var descriptions

meta2 <- 
    ff_meta %>% 
        filter(scale %in% scales_once |
            (scale %in% scales_multiple &
             wave == "Year 15")) %>%
        select(topics,
             subtopics,
             scale,
             new_name, 
             varlab,
             type,
             source, 
             respondent, 
             wave,
             starts_with("label"))

rio::export(meta2, here::here("data","meta2.csv"))

variables <- rbind(
    rio::import(here("data","meta1.csv")),
    rio::import(here("data","meta2.csv"))
)


variables <- variables %>% 
    select(-c(paste0("label",11:30)))

rio::export(variables, here::here("data","variables.xlsx"))
```

```{r}
#use variable names to subset the dataset

ff <- haven::read_sav(here("data","FF_allwaves_2020.sav")) %>% 
    clean_names()

ff_sub <- ff %>% 
    select(idnum, #identifier 
           all_of(scale_vars$new_name)) %>%  #our selected variables
    as_tibble()

#assigning attributes to variables

var <-colnames(ff_sub)[-1]
atr <- c("wave", "respondent", "scale", "survey")    #can add more meta-data info that we want readily accesible


for(i in seq_along(var)){
    for(j in seq_along(atr))
    {
        attr(ff_sub[[i+1]], atr[j]) <- as.character(
            scale_vars %>% 
                filter(new_name == var[i]) %>% 
                select(atr[j]))
        
    }
}

#now, variable description can be seen as attributes(data$var_name)
#attributes(ff_sub$k6d45)

#export(ff_sub, here("data","ff_sub.rda"))
```


```{r, eval = FALSE}
ff_sub <- rio::import(here("data","ff_sub.rda"))
```

Documentation on scales:

- CBCL 
1. Collected 4 times, in year 3, 5, 9, and 15. 
2. Information in year 3 (childcare provider-64Q and primary caregiver-55Q), 5 (primary caregiver - 53Q and mother - 32Q), 9 (primary caregiver - 111 Qs), 15 (primary caregiver - 34 Qs)
3. 3-point Likert scale
4. Have seven sub-scales -- only cleaning for year 15
5. Some similar items over different years
6. Guide for year 15: pg 48

- SDQ: 
1. Collected once in year 9 
2. 14 items 
3. 4-point Likert scale: 0-3 likert scale: not at all true, somewhat true, mostly true, very true 
4. Two sub-scales: internalizing and externalizing behavior
5. Guide for year9: pg 54

- Trouble at schoole: 
1. Collected once in year 15. 
2. 4 items. 
3. 3-point Likert scale (1-3: Never, sometimes, often) an 7 = NA/Homeschooled 
4. Guide for year 15: pg 58
5. Composite score calculated

- Peer bullying
1. Collected twice year 9 and year 15
2. 4 items 
3. 4-point Likert scale (0-3)  
4. Guide for year 15: pg 56
5. Composite score calculated

- Teen Tobacco Use: 
1. Collected once in year 15
2. 6 items 
3. some are yes/no, some on frequency 
4. Guide for year 15: pg 35
5. Composite score not calculated

I'll find composite scores for the 4 scales (except teen tobacco use), and keep the items from teen tobacco use in factor form as is. 

Each subscale has three steps -
1. Count number of missing
2. Decision based on missing values - scale total can't be calculated, or calculate scale total
3. Scale total - either sum or mean of responses

Will use a function for each of these and call them to make scale/sub-scale totals

```{r}
#all negative are NA. 7 means home-schooled and is also NA
ff_sub_clean <- ff_sub %>%
  mutate(
    across(
      .cols = everything(),
      .fns = function(x) ifelse(x<0 | x == 7 , NA, x)
    )
  ) 

#doing the above step stripped attributes. Assigning attributes again
#give back the attributes
attributes(ff_sub_clean) <- attributes(ff_sub)

#one column is continuous, doing sjlabelled overwrites its values
for (i in c(2:58, 60:63)){
  attributes(ff_sub_clean[[i]]) <- attributes(ff_sub[[i]])
}

ff_sub_clean <- ff_sub_clean %>% 
  sjlabelled::as_label()
```


```{r}
#Functions - find missing values for each sub-scale
count_missing <- function(vars){
  temp <- ff_sub_clean %>% 
  select(idnum,
         vars) %>% 
  mutate(
    n_missing = apply(., 1, function(x) sum(is.na(x)))
  )  
  
  temp
}

#Calculate composite score
subscale_summary <- function(df, criteria = 0, operation = mean, adjust = 0){
  temp <- df %>% 
  pivot_longer(
    cols = -c('idnum','n_missing'),
    names_to = 'var',
    values_to = 'val'
  ) %>% 
  mutate(
    val = parse_number(as.character(val))
  ) %>%
  group_by(idnum, n_missing) %>% 
  summarise(
    score = ifelse(n_missing > as.numeric(criteria),
                   NA,
                   `operation`(val, na.rm = TRUE)
                   )
  ) %>% 
  unique()
  
  temp$score - adjust
}

#Combine the two functions
subscale_score <- function(vars, criteria = 0, operation = mean, adjust = 0){
  temp <- count_missing(vars)
  subscale_summary(temp, criteria, operation, adjust)
}

```

```{r}
#subscale variables

#vars argument for the function
subscales <- list(
sdq9_externalizing = paste0("k5g2",c(letters[c(2,4,6,8,13,14)])),
sdq9_internalizing = paste0("k5g2",c(letters[c(1,3,5,7,9:12)])),
cbcl15_aggresive = paste0("p6b",c(35, 37:39, 41:45, 57, 59)),
cbcl15_anxious = paste0("p6b",c(36, 40, 52:54, 68)),
cbcl15_attention = paste0("p6b",c(46:48)),
cbcl15_rulebreaking = paste0("p6b",c(49:51, 60:64, 67)),
cbcl15_socialproblem = "p6b56",
cbcl15_thoughtproblem = "p6b55",
cbcl15_withdrawn = paste0("p6b",c(65:66)),
# cbcl15_internalizing = cbcl15_axious + cbcl_withdrawn,
# clcb15_externalizing = cbcl15_aggresive + cbcl15_rulebreaking
pb15 = paste0("k6b32",c(letters[c(1:2,5:6)])),
tas15 = paste0("k6b21", c(letters[c(1:4)]))
)

#criteria - 2 missing allowed for externalizing and internalizing and 0 for others

subscale_criteria <- c(2,2, rep(0,9))

#adjust - cbcl mean score has to be reduced by 1

subscale_adjust <- c(0,0, rep(1,7), 0, 0)
```


```{r}
#making a dataframe with composite scores and responses for Tobacco Use Survey
subscale_df <- data.frame(
  idnum = c(1:4898)
)

for(i in seq_along(subscales)){
   x <- subscale_score(subscales[[i]], 
                 subscale_criteria[i],
                 mean,
                 subscale_adjust[i])
   
   subscale_df <- cbind(subscale_df, x)
}

#fix col names
colnames(subscale_df)[2:12] <- attributes(subscales)$names

```

```{r}
#Items from teen tobacco use
ttu15 <- paste0("k6d",c(40, 44, 45)) #k6d41-43 has almost all missing values as most respondents skipped it. Therefore not including

#this is the clean df that can be used for modelling
subscale_df <- ff_sub_clean %>%
  select(idnum, ttu15) %>% 
  mutate(
    idnum = as.numeric(idnum)
  ) %>% 
  right_join(subscale_df)

  
```


